manuscript_modelling_and_plots_takeovers

### 2024-08-30 manuscript_and_modelling_plots_takeovers.Rmd
new code for plotting and model fitting with truncation

pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "stat_grouped", stat = "mean", group = "n_back")

pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

d <- pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "hist", group = "ttc_criticality_factor")

ggplot(d[["data"]], mapping = aes(value)) +
  geom_histogram(aes(fill = as.factor(is_y))) +
  facet_wrap(~ rep_id)

# modelling average deceleration with regularising priors. Random effects structure is maximal. Distribution family is Gaussian.
mean_decel_gaussian_reg_priors_maximal <- brm(data = response.vehicle.eye %>%
                                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
                                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                                             dplyr::mutate(ppid = as.factor(ppid)),
                                           family = gaussian(),
                                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept"),
                                                     prior(student_t(3, 0, 2.5), class = "b")),
                                           bf(mean_decel ~ n_back * ttc_criticality_factor + (n_back * ttc_criticality_factor | ppid)),
                                           control = list(adapt_delta = 0.99999999999),
                                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13,
                                           file = here::here("takeover_performance_cogload_time_budget/Models/mean_decel_gaussian_reg_priors_maximal"))
# summarising model parameters 
print(summary(mean_decel_gaussian_reg_priors_maximal), digits = 5)
describe_posterior(mean_decel_gaussian_reg_priors_maximal, centrality = "mean")

pp_check(mean_decel_gaussian_reg_priors_maximal, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
pp_check(mean_decel_gaussian_reg_priors_maximal, type = "stat_grouped", stat = "mean", group = "n_back")

pp_check(mean_decel_gaussian_reg_priors_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(mean_decel_gaussian_reg_priors_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "hist", group = "ttc_criticality_factor")
```

@@

```{r}
# summarising model parameters 
print(summary(mean_decel_skewnormal_reg_priors_maximal), digits = 5)
describe_posterior(mean_decel_skewnormal_reg_priors_maximal , centrality = "mean")

@@

# posterior predictive checks
pp_check(sd_yaw_automation_gaussian_maximal, type = "stat_grouped", stat = "mean", group = "n_back")
pp_check(sd_yaw_automation_gaussian_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
```

@@

@@

@@

pp_check(sd_yaw_manual_gaussian_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(sd_yaw_manual_gaussian_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)
```

@@

```{r}
print(summary(sd_yaw_manual_gaussian_maximal), digits = 5)
describe_posterior(sd_yaw_manual_gaussian_maximal, centrality = "mean")

@@

# predicted mean SD yaw  implied by the model as a function of TTC # setting re_formula = NA tells the function to not use any random effects - here we plot the n-back specific distributions 
predicted_sd_yaw_manual_ttc <- sd_yaw_manual_gaussian_maximal %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3), as.factor(5)),
                                    n_back = FALSE),
              re_formula = NA)

@@

# margainal effect of TTC for SD of yaw
sd_yaw_manual_MER_TTC <- sd_yaw_manual_gaussian_maximal %>% 
  emmeans(~ ttc_criticality_factor,
          at = list(n_back = FALSE),
          epred = TRUE, re_formula = NA) %>% 
  contrast(method = "revpairwise") %>% 
  gather_emmeans_draws()

@@

## predicted means for the effect of n-back on SD yaw
predicted_sd_yaw_manual_nback <- sd_yaw_manual_gaussian_maximal %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3)),
                                    n_back = c(FALSE, TRUE)),
              re_formula = NA)

@@

# marginal effect of N-back of SD of yaw during manual driving
sd_yaw_manual_MER_nback <- sd_yaw_manual_gaussian_maximal %>% 
  emmeans(~ n_back,
          at = list(ttc_criticality_factor = as.factor(3)),
          epred = TRUE, re_formula = NA) %>% 
  contrast(method = "revpairwise") %>% 
  gather_emmeans_draws()

>>>>

mean_decel_trunc_gaussian_reg_priors_maximal <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/mean_decel_trunc_skewnormal_reg_priors_maximal.rds"))

# modelling average deceleration with regularising priors. Random effects structure is maximal. Distribution family is skew-normal. 
mean_decel_trunc_gaussian_reg_priors_maximal <- brm(data = response.vehicle.eye %>%
                                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
                                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                                             dplyr::mutate(ppid = as.factor(ppid)),
                                           family = gaussian(),
                                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept"),
                                                     prior(student_t(3, 0, 2.5), class = "b")),
                                           bf(mean_decel | trunc(ub = 0) ~ n_back * ttc_criticality_factor + (n_back * ttc_criticality_factor | ppid)),
                                           control = list(adapt_delta = 0.99999999999),
                                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13,
                                           file = here::here("takeover_performance_cogload_time_budget/Models/mean_decel_trunc_gaussian_reg_priors_maximal"))

mean_decel_gaussian_reg_priors_maximal <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/mean_decel_gaussian_reg_priors_maximal.rds"))

# modelling average deceleration with regularising priors. Random effects structure is maximal. Distribution family is Gaussian.
mean_decel_gaussian_reg_priors_maximal <- brm(data = response.vehicle.eye %>%
                                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
                                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                                             dplyr::mutate(ppid = as.factor(ppid)),
                                           family = gaussian(),
                                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept"),
                                                     prior(student_t(3, 0, 2.5), class = "b")),
                                           bf(mean_decel ~ n_back * ttc_criticality_factor + (n_back * ttc_criticality_factor | ppid)),
                                           control = list(adapt_delta = 0.99999999999),
                                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13,
                                           file = here::here("takeover_performance_cogload_time_budget/Models/mean_decel_gaussian_reg_priors_maximal"))
```

@@

```{r}
# summarising model parameters 
print(summary(mean_decel_skewnormal_reg_priors_maximal), digits = 5)
describe_posterior(mean_decel_skewnormal_reg_priors_maximal, centrality = "mean")

@@

sd_yaw_automation_lognormal_maximal <- brm(data = response.vehicle.eye %>%
                                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, collision == 0, ho.rt > .15) %>%
                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                             dplyr::mutate(ppid = as.factor(ppid)),
                           family = lognormal(),
                           sd_yaw_automation ~ n_back + (n_back | ppid),
                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept"),
                                     prior(student_t(3, 0, 2.5), class = "b")),
                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13, 
                                                   file = here::here("takeover_performance_cogload_time_budget/Models/sd_yaw_automation_lognormal_maximal"))

@@

sd_yaw_automation_MER_nback  %>% mean_hdi()

## Log normal model summary

# model summary
print(summary(sd_yaw_automation_lognormal_maximal), digits = 5)
describe_posterior(sd_yaw_automation_lognormal_maximal, centrality = "mean")

# average predicted SD of yaw during automation with and without n-back
predicted_sd_yaw_automation_nback <- sd_yaw_automation_lognormal_maximal %>% 
  epred_draws(newdata = expand_grid(n_back = c(TRUE, FALSE)),
              re_formula = NA)

# average marginal effect of n-back
sd_yaw_automation_MER_nback <- sd_yaw_automation_lognormal_maximal %>% 
  emmeans(~ n_back,
          epred = TRUE, re_formula = NA) %>% 
  contrast(method = "revpairwise") %>% 
  gather_emmeans_draws()

@@

sd_yaw_manual_trunc_gaussian_maximal <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/sd_yaw_manual_trunc_gaussian_maximal.rds"))

# initial model of sd_yaw
sd_yaw_manual_trunc_gaussian_maximal <- brm(data = response.vehicle.eye %>%
                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, collision == 0, ho.rt > .15) %>%
                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                             dplyr::mutate(ppid = as.factor(ppid)),
                           family = gaussian(),
                           bf(sd_yaw_manual | trunc(lb = 0) ~ n_back * ttc_criticality_factor + (n_back * ttc_criticality_factor| ppid)),
                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept"),
                                     prior(student_t(3, 0, 2.5), class = "b")),
                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13, 
                           control = list(adapt_delta = 0.95),
                           file = here::here("takeover_performance_cogload_time_budget/Models/sd_yaw_manual_trunc_gaussian_maximal"))

@@

pp_check(sd_yaw_manual_gaussian_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(sd_yaw_manual_gaussian_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

pp_check(sd_yaw_manual_trunc_gaussian_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(sd_yaw_manual_trunc_gaussian_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)
```

@@

```{r}
print(summary(sd_yaw_manual_trunc_gaussian_maximal), digits = 5)
describe_posterior(sd_yaw_manual_trunc_gaussian_maximal, centrality = "mean")

@@

# predicted mean SD yaw  implied by the model as a function of TTC # setting re_formula = NA tells the function to not use any random effects - here we plot the n-back specific distributions 
predicted_sd_yaw_manual_ttc <- sd_yaw_manual_trunc_gaussian_maximal %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3), as.factor(5)),
                                    n_back = FALSE),
              re_formula = NA)

@@

# margainal effect of TTC for SD of yaw
sd_yaw_manual_MER_TTC <- sd_yaw_manual_trunc_gaussian_maximal %>% 
  emmeans(~ ttc_criticality_factor,
          at = list(n_back = FALSE),
          epred = TRUE, re_formula = NA) %>% 
  contrast(method = "revpairwise") %>% 
  gather_emmeans_draws()

@@

## predicted means for the effect of n-back on SD yaw
predicted_sd_yaw_manual_nback <- sd_yaw_manual_trunc_gaussian_maximal %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3)),
                                    n_back = c(FALSE, TRUE)),
              re_formula = NA)

@@

# marginal effect of N-back of SD of yaw during manual driving
sd_yaw_manual_MER_nback <- sd_yaw_manual_trunc_gaussian_maximal %>% 
  emmeans(~ n_back,
          at = list(ttc_criticality_factor = as.factor(3)),
          epred = TRUE, re_formula = NA) %>% 
  contrast(method = "revpairwise") %>% 
  gather_emmeans_draws()
